---
title: "Assignment 3"
author: "Navneeth I U"
date: "28 September 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, message=FALSE, warning=FALSE}
sensor_data = read.csv("C://Users//Administrator//Documents//EDA//CSV//data_pca.csv")
library(dplyr)
library(corrplot)
dim(sensor_data)
```


##### Principal Components 
```{r, message=FALSE, warning=FALSE}
sensor_pca = prcomp(sensor_data, scale. = T)
```


##### Scree diagram 
```{r, message=FALSE, warning=FALSE}
screeplot(sensor_pca, main = "", type = "l")
title("Scree Plot", xlab = "Principal Components")
```


##### Output variables from prcomp function 
```{r, message=FALSE, warning=FALSE}
names(sensor_pca)
```

##### We use the name function to obtain output variables
##### sdev - This variable gives the standard deviation of the PC's or square root of eigen values 
##### rotation - This output variable gives the eigen vectors 
##### center  - This variable gives the mean of the mean of the input variables 
##### scale - This variable gives the standard deviation of the input variables 
##### x - This variable gives the principle components 

#### Eigen values
```{R, message=FALSE, warning=FALSE}
sensor_eigen_values = (sensor_pca$sdev*sensor_pca$sdev)
sensor_eigen_values
```


##### Cumulative variance 
```{r, message=FALSE, warning=FALSE}
cum_var = cumsum(sensor_eigen_values/length(sensor_eigen_values)*100)
cum_var
a = matrix(c(0,50,80,80),2,2, byrow = FALSE)
plot(cum_var, type = "l", col = "blue", main = "cumulative variance by PC's", 
     ylab = "% Variance of PCs", xlab = "Principal Components") + lines(a, col = "red")
```



##### From the above plot we can identify that 2 PC's csn retain 80% of variance of the data



##### A biplot uses points to represent the scores of the observations on the principal components, and it uses vectors to represent the coefficients of the variables on the principal components. A vector points in the direction which is most like the variable represented by the vector.
```{r, message=FALSE, warning=FALSE}
biplot(sensor_pca, choices = 1:2, scale = 1)
```


##### Airquality data 
```{r, message=FALSE, warning=FALSE}
head(airquality)
```


##### Identifying the percentage of null values 
```{r, message=FALSE, warning=FALSE}
percent_nulls = colSums(is.na(airquality))/NROW(airquality)*100
percent_nulls
```


##### aggr plot using "VIM" package 
```{r, message=FALSE, warning=FALSE}
library(VIM)
aggr(airquality, delimiter = NULL, plot = T)
```



##### Data imputation using "mice" package 
##### The mice package in R, helps you imputing missing values with plausible data values. These plausible values are drawn from a distribution specifically designed for each missing datapoint.
```{r, message=FALSE, warning=FALSE}
library(mice)
data_imp = mice(airquality, m=5)
```

